DAOS and Friends: A Proposal for\\an Exascale Storage System

### 2016-07-26 paper.tex
added reviews and completed final version edits.

\documentclass[conference]{IEEEtran} \pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\usepackage{hyperref}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

@@

\begin{document}

%\title{Fast Forwarding Storage for Extreme Scale Platforms and Big Data Applications}
\title{Lessons Learned from the Fast Forward Storage and I/O Project}

\author{
%\IEEEauthorblockN{Blinded for review},\IEEEauthorblockA{}
\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Lawrence Berkeley National Laboratory koziol@lbl.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Segate Government Solutions John.Bent@seagategov.com}
\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
}
\maketitle

\begin{abstract}
%The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
%project is going to have significant impact on storage systems design within
%and beyond the HPC community. With phase two of the project starting, it is an
%excellent opportunity to explore the complete design and how it will address
%the needs of extreme scale platforms.  This paper examines the overall design.
%Each layer of the proposed stack is discussed in some detail along with
%cross-cutting topics that impact multiple layers, such as transactions and
%metadata management. We also describe the proposed features to better support
%some classes of big data applications.
%
%This paper not only provides a timely summary of important aspects of the
%design specifications but also captures the underlying reasoning that is not
%available elsewhere. With this information, we encourage the broader community
%to understand the design, intent, and future directions to foster discussion
%guiding phase two and the ultimate production storage stack based on this work.
%An initial performance evaluation of the early prototype implementation is also
%provided to validate the presented design.

The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase two of the project starting, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines each layer of the
proposed stack in some detail along with cross-cutting topics, such as
transactions and metadata management.

@@

To address these challenges, a joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase completed in 2014 with a second phase
currently getting underway. The core focus of the first phase was basic
functionality and design. While an idealized potential system would be the
perfect target architecture, the reality of budgets has tempered many of the
decisions. For example, extensive availability of NVRAM or SSDs on all of the
compute nodes is currently not economically feasible limiting some of the
potential design choices.  With this in mind, the second phase will refine this
design incorporating fault recovery and other features missing from the first
phase.

The complete design seeks to offer high availability, byte-granular,
multi-version concurrency control. Through the use of a copy-on-write style
mechanism reducing storage space and writing time requirements, multiple
versions of an object can be stored efficiently. By assuming the client
interface will be through an IO library, a more complicated interface offering
richer functionality can be incorporated while requiring only minimal end-user
code changes.  Managing most data access in a platform-local layer rather than
requiring writing to centralized storage will better support the performance
and energy requirements of extreme scale application compositions.

Overall, the architecture shifts from the idea of files and directories to
containers of objects. This shift avoids the bottlenecks related to the POSIX
files and directories structure such as the serialization of file creation, the
limitation and impact of the number of files in a directory, and the limited
semantics of a byte stream. Instead, the new interface focuses on high-level
data models and their properties and relationships. This concept permeates the
entire IO stack.

In addition to addressing the traditional scientific workload, this project
seeks to expand functionality to better support Big Data type applications. The
key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
Map-Reduce systems. Key system features are introduced to efficiently support
these computing models in addition to the typically bursty IO loads of more
traditional HPC applications.

@@

At a more detailed view, the various layers of the IO stack each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is comprised of generally a high level IO library, such as the
demonstration HDF5 library~\cite{hdf5} and a more complex API for accessing the
lower level components. This layer is in dark blue. Becuase the system supports
more complex architectures and supports richer functionality, hiding this
complexity behind a user-friendly API is the intent. It is possible to 
access the storage stack through the more complex API, but the additional
requirements beyond standard POSIX calls will prompt most users to use an IO
library.  This layer incorporates the necessary features for ACGs from a
end-user's perspective.

@@

The Distributed Asynchronous Object Storage (DAOS) layer serves as the
persistent storage interface and translation layer between the user-visible
object model and the requirements of the underlying storage infrastructure. 
Transactions work a bit differently at this layer and are called epochs to
distinguish them. DAOS is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance and usability penalty of
manipulating the shared, persistent storage array directly.

@@

Along with analysis of the published design documents, a discussion of the
design philosophy representing the overall intent is presented. This
information represents information that may or may not have been written down,
but is the intent of ultimate system.  This information was gathered through
interviews with the core FFSIO team members. These ideas are presented to give
a fuller picture of where the project is going rather than dwelling on any
limitations of the published documents. This is most important to illustrate
how different concepts will work across layers since that information is spread
across multiple documents and may lack a cohesive overall view. Previously,
a poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
been presented.

The rest of the paper is organized as follows. An overview of related work is
presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but discussed more fully in the cross-cutting section. The VOSD
layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in
parallel file systems. Ceph's unique striping approach uses pseudo-random
numbers with a known seed eliminating the need for the metadata service to
track where each stripe in a parallel file is placed. Ceph's strengths are in
providing good perforamnce and scalability with the ability to handle failures
and deploying new storage adapting the system in a live environment. However,
this failure handling advantage was shown~\cite{wang:2013:ceph} to limit peak
performance more than other systems like Lustre.  More recently, some of these
limitations have begun to be addressed by the Ceph team, but no new evaluations
have been performed to determine if these changes close the gap sufficiently to
address the extreme HPC performance needs.

@@

HDF5~\cite{hdf5} has a versatile data model offering complex data objects
and metadata. Its information set is a collection of datasets, groups,
datatypes and metadata objects. The data model defines mechanisms for
creating associations between various information items. The main
conceptual components for data stored in HDF5 are described below.

@@

The Virtual Object Layer is an abstraction mechanism internal to the HDF5
library~\cite{hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
just below the public API. The VOL exports an interface that allows writing
plugins for HDF5 enabling developers to handle data in ways other than writing
to storage in an HDF5 format.  Plugin writers provide an implementation for a
set of functions and are trusted to provide the proper semantics for the new
environment. For example, data staging could be implemented in the VOL layer by
replacing writing to disk in the HDF5 format to sending data to a data staging
area using some messaging mechanism.

For this project, rather than the default writing to disk in the HDF5 format,
the VOL is used to interact with the IOD layer and the different concepts it
offers without requiring all of the functionality be exposed to users. For
example, the containers and objects concept is transparently mapped to the
files and datasets existing HDF5 users are familiar with.  This reduces the
difficulty porting applications to the new IO stack.

@@

Since HDF5 has traditionally offered an interface focused on files and the
internal data types, such as datasets, these concepts must be mapped onto the
proposed FFSIO data storage concepts.  This mapping is shown in
Table~\ref{tab:mapping}

\begin{table}[ht]
%    \vspace{-0.10in}
    \centering
    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
    \bigskip
%    \vspace{-0.15in}

    \begin{tabular}{|r|r|}
\hline
HDF5 & FFSIO\\
\hline
file & container \\
dataset & array \\
group & key-value store \\
attribute(s) & key-value store \\
\hline
    \end{tabular}
    \label{tab:mapping}
\end{table}

In Section~\ref{sec:iod} below, the FFSIO types are described in more detail.

@@

The core idea for IOD is to provide a way to manage IO load that is
separate from the compute nodes and the storage array. Communication intensive
activities, such as data rearrangement, can be moved to the IOD layer
reducing the number of participants and message count. IOD has three main
purposes. First, the burst buffers work as a fast cache absorbing write
operations that then trickles out to the central storage array. It can also
be used to retrieve objects from the central storage array for more efficient
read operations and offers data filtering to make client reads more efficient.
Second, it offers the transaction mechanism for controlling data set visibility
and to manage faults that could expose an incomplete or corrupt data set to
users. These transactions are local to the IOD layer until persisted to the
DAOS layer eliminating the need for burdening the persistent storage with
transient data.  Third, data processing operations can be placed in the IOD.
These operations are intended to offer functionality like data rearrangement
and filtering prior to data reaching the central storage array.

@@

\subsubsection{Container}
As mentioned above, the concept of a container is similar to that of a file in
a traditional file system. However, rather than being in a directory structure,
each container essentially is stored in a hash-space allowing direct access to
any container without regard to the current organizational context of the
file system. For example, there is no need to navigate a directory hierarchy
to name a particular container.

@@

\subsubsection{Key-Value Store}
This is the base type for the container. Since the container represents
something akin to HDF5 files, everything is stored within a hierarchical
namespace. The root namespace is represented by the base key-value store and
contains a list of all of the objects for this portion of the namespace as well
as additional key-value objects representing sub-groups for the hierarchy. Each
of those key-value store objects works identically.  Attributes are stored in a
key-value store object, but use the
multi-dimensional array and blob objects to store the values for the
attributes.

\subsubsection{Multi-Dimensional Array}
By treating arrays as a special case separate from blobs, additional
opportunities are enabled. For example, by knowing that an object is an array,
proper slicing of that array onto IO nodes can be done without involving higher
levels of the IO stack.

\subsubsection{Blob}
All other data is stored as a stream-of-bytes without regard to the actual
data type.

\subsection{Multi-Dimensional Array Data Distribution}

For both IO performance and to aid in analysis and other data processing, the
multi-dimensional array object can be split across multiple IO nodes. Each
piece of this array is called a {\em shard}.  The idea of sharding is to store
a logically complete portion of a data set on a single storage target. This is
similar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
the data in the default or some other structured way as well as ``re-sharding''
based on application needs. For example, reordering the data so that a
different dimension is the ``fast'' dimension may greatly improve the
performance of a subsequent data analytics task. A common scenario where this
is common is a Fortran code (column-major) writes data for a C code (row-major)
to analyze.  The IOD API supports the following sharding strategies:

\begin{itemize}
\item
  \textbf{contiguous}. fixed chunking, distributed in a round-robin
  fashion across the IO nodes.
\item
  \textbf{chunked}. same as above but with irregular (sparse) chunking.
\item
  \textbf{user-defined}. either contiguous or chunked, but user specifies
  where to place each individual shard.
\end{itemize}

It is possible to request the transformation of an object's physical
layout to other formats resulting in multiple copies of the same objects in
multiple formats if desired. Also, the user can pre-fetch objects from
the storage cluster into the IO nodes or read them directly from the
storage cluster. At the semantic level (HDF5), indices can be created
for datasets resulting in being able to read through an index instead of
directly from the base array.

@@

The Distributed Asynchronous Object Storage layer serves as the traditional
parallel file system interface layer for the storage devices. This is the
consistent, global view of the underlying devices represented in this stack
by the VOSD layer.

This is the layer where the container/object model is translated into the
physical storage requirements dictated by the physical storage underneath (the
VOSD layer). The two key design elements of this layer are the handling of
epochs and the mapping of containers and objects to the underlying storage.

@@

In the subsections that follow, we examine transactions and epochs, metadata
management, function and analysis shipping, and the arbitrary connected graphs
support.

@@

@@

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
This is a hybrid approach to metadata management that is halfway between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

@@

\begin{figure*}[htbp!]
\centering
%\vspace{-0.10in}
\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
%\vspace{-0.10in}
\caption{Functionality Demonstration Validation for Number of Hosts}
\label{fig:eval-hosts}
%\vspace{-0.05in}
\end{figure*}

\begin{figure*}[htbp!]
\centering
%\vspace{-0.10in}
\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
%\vspace{-0.10in}
\caption{Functionality Demonstration Validation for Data Sizes}
\label{fig:eval-size}
%\vspace{-0.05in}
\end{figure*}

@@

All of the tests are performed on
a blinded for review machine.
%the Buffy Cray XC30-AC at LANL.
It consists
of 64 compute nodes each with dual, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz
and 64 GB RAM.  The interconnect is Cray Aries. There are 14 IO nodes
consisting of single socket, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz with 32
GB of RAM.  There are also a metadata, login, and 2 boot nodes.  The storage
array has disk and SSD partitions. One is a DDN Lustre system with 192 TB disk
usable with a minimum of 5 GB/sec performance. The rest are SSDs consisting of
an EMC flash array connected via FDR InfiniBand with 22 TB and 48 GB/sec write
performance.

@@

With the overall stack design a prototype implementation complete, refinements,
such as fault detection and recovery, can be designed and tested.  These and
other activities for phase 2 will ultimately generate what is likely to be the
next generation storage stack for extreme scale platforms.

\section{Acknowledgements}

%\includegraphics[scale=0.07]{logos/doe_logo}
%\includegraphics[scale=0.30]{logos/snl_logo}
%\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

>>>>

\documentclass[conference]{IEEEtran} \pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\IEEEoverridecommandlockouts
\IEEEpubid{\makebox[\columnwidth]{978-1-4673-8815-3/16/\$31.00 ©2016 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
\begin{document}

\title{DAOS and Friends: A Proposal for\\an Exascale Storage System}

\author{
\IEEEauthorblockN{Jay Lofstead\IEEEauthorrefmark{1}, Ivo Jimenez\IEEEauthorrefmark{2}, Carlos Maltzahn\IEEEauthorrefmark{2}, Quincey Koziol\IEEEauthorrefmark{3}, John Bent\IEEEauthorrefmark{4}, Eric Barton\IEEEauthorrefmark{5}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Sandia National Laboratories gflofst@sandia.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{2}University of California, Santa Cruz ivo@cs.ucsc.edu, carlosm@soe.ucsc.edu}
\IEEEauthorblockA{\IEEEauthorrefmark{3}Lawrence Berkeley National Laboratory koziol@lbl.gov}
\IEEEauthorblockA{\IEEEauthorrefmark{4}Segate Government Solutions John.Bent@seagategov.com}
\IEEEauthorblockA{\IEEEauthorrefmark{5}Intel eric.barton@intel.com}
}
\maketitle

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase two of the project starting, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines each layer of the
proposed stack in some detail along with cross-cutting topics, such as
transactions and metadata management.

@@

A joint effort between the US Department of
Energy's Office of Advanced Simulation and Computing and Advanced Scientific
Computing Research commissioned a project to develop a design and prototype for
an IO stack suitable for the extreme scale environment. It will be referred to
as the Fast Forward Storage and IO (FFSIO) project. This is a joint effort led
by Lawrence Livermore National Laboratory, with the DOE Data Management Nexus
leads Rob Ross and Gary Grider as coordinators and contract lead Mark Gary. The
participating labs are LLNL, SNL, LANL, ORNL, PNL, LBNL, and ANL.  Additional
industrial partners contracted include the Intel Lustre team, EMC, DDN, and the
HDF Group. This team has developed a specification
set~\cite{fastforward:2014:docs} for a future IO stack to address the
identified challenges. The first phase completed in 2014 with a second phase
underway. The first phase focused primarily on basic
functionality and design. While an idealized potential system would be the
perfect target architecture, the reality of budgets has tempered many of the
decisions. For example, extensive availability of NVRAM or SSDs on all of the
compute nodes is currently not economically feasible limiting some of the
potential design choices.  With this in mind, the second phase is 
incorporating fault recovery and other missing features.

The complete design seeks to offer high availability, byte-granular,
multi-version concurrency control. Multiple versions of an object are stored
efficiently by using a copy-on-write mechanism. By assuming the client
interface will be through an IO library, a more complicated interface offering
richer functionality can be incorporated while requiring only minimal end-user
code changes.  Managing most data access in a platform-local layer rather than
requiring writing to centralized storage will better support the performance
and energy requirements of extreme scale application compositions.

Overall, the architecture shifts from the idea of files and directories to
containers of objects. This shift avoids the bottlenecks related to the POSIX
files and directories structure such as file creation serialization, the
file count in a directory limitation and impact, and the limited
semantics of a byte stream. Instead, the new interface focuses on high-level
data models and their properties and relationships. This concept permeates the
entire IO stack.

In addition to addressing the traditional scientific workload, this project
seeks to expand functionality to better support Big Data type applications. The
key idea is to support Arbitrary Connected Graphs (ACGs) such as those used in
Map-Reduce systems. Key system features are introduced to efficiently support
these computing models in addition to the typically bursty IO loads of more
traditional HPC applications. These features are not discussed for space
reasons.

@@

The IO stack layers each contribute
different functionality.  The architecture (Figure~\ref{fig:arch-mapping})
incorporates five layers, some of which have potentially optional components.
The top layer is comprised of generally a high level IO library, such as the
demonstration HDF5 library~\cite{folk:2011:hdf5} and a more complex API for accessing the
lower level components. This layer is in dark blue. Becuase the system supports
more complex architectures and supports richer functionality, hiding this
complexity behind a user-friendly API is the intent. It is possible to 
access the storage stack through the more complex API, but the additional
requirements beyond standard POSIX calls will prompt most users to use an IO
library.  This layer incorporates the necessary features for ACGs from a
end-user's perspective.

@@

The Distributed Application Object Storage (DAOS) layer serves as the
persistent storage interface and translation layer between the user-visible
object model and the requirements of the underlying storage infrastructure. 
Transactions work a bit differently at this layer and are called epochs to
distinguish them. DAOS is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance and usability penalty of
manipulating the shared, persistent storage array directly.

@@

This paper presents an analysis of the published design documents along with a
discussion of the design philosophy representing the overall intent. The design
philospohy conveys information that may or may not have been written down, but
aids understanding for the total design.  This information was gathered through
interviews with the core FFSIO team members. These ideas are presented to
reveal the project future rather than dwelling on any limitations of the
published designs. This is most important to illustrate how different concepts
will work across layers since that information is spread across multiple
documents and may lack a cohesive overall view. Previously, a
poster~\cite{lofstead:2014:ffsio-poster} and a more detailed evaluation of
consistency and fault tolerance~\cite{lofstead:2014:ffsio-consistency} have
been published.

With DAOS and friends being groomed as the next generation for the Lustre
parallel file system, the information and analysis presented here can help
users determine how to adapt their thinking about storage as well as hopefully
influence the second phase.

The key contributions of this paper are the syntesized architectural overview
and the discussion and analysis motivating the architectural descisions. With
this information, community members can be better informed about an important
future storage system as well as potentially influence the evolution of the
design from prototype into production system.

The rest of the paper is organized as follows. An overview of related work is
presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array.  This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration.
Section~\ref{sec:iof} briefly discusses the motivation and proposal for the IO
forwarding layer.  Section~\ref{sec:iod} describes the IO Dispatcher layer and
the broad functionality it offers.  This will detail the pieces of the layer
that are potentially optional and mention the cross-cutting features discussed
in a later, cross-cutting section.  Section~\ref{sec:daos} discusses how the
DAOS layer functions.  As with the IOD layer, the cross-cutting features will
be mentioned, but discussed more fully in the cross-cutting section.  The VOSD
layer is discussed in Section~\ref{sec:vosd}.  In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage.  Next is an exploration of cross-cutting features like transactions
and metadata management in Section~\ref{sec:broader}.  Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.  This shows that the prototype system based on
the proposed design can function.  Section~\ref{sec:conclusion} concludes the
paper with a summary of the broad issues.

@@

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in
parallel file systems. Ceph's unique striping approach uses pseudo-random
numbers with a known seed eliminating the need for the metadata service to
track where each piece in a striped file is placed. Ceph's strengths are in
providing good perforamnce and scalability with the ability to handle failures
and deploying new storage adapting the system in a live environment. However,
this failure handling advantage was shown~\cite{wang:2013:ceph} to limit peak
performance more than other systems like Lustre.  More recently, some of these
limitations have begun to be addressed by the Ceph team, but no new evaluations
have been performed to determine if these changes close the gap sufficiently to
address the extreme HPC performance needs.

@@

HDF5~\cite{folk:2011:hdf5} has a versatile data model offering complex data objects
and metadata. Its information set is a collection of datasets, groups,
datatypes and metadata objects. The data model defines mechanisms for
creating associations between various information items. The main
conceptual components for data stored in HDF5 are described below.

@@

The Virtual Object Layer is an abstraction mechanism internal to the HDF5
library~\cite{folk:2011:hdf5}. As shown in Figure~\ref{fig:vol-arch} it is implemented
just below the public API. The VOL exports an interface that allows writing
plugins for HDF5 enabling developers to handle data in ways other than writing
to storage in an HDF5 format.  Plugin writers provide an implementation for a
set of functions and are trusted to provide the proper semantics for the new
environment. For example, data staging could be implemented in the VOL layer by
replacing writing to disk in the HDF5 format to sending data to a data staging
area using some messaging mechanism.

For this project, rather than the default writing to disk in the HDF5 format,
the VOL is used to interact with the IOD layer and the different concepts it
offers without requiring all of the functionality be exposed to users. For
example, the containers and objects concept is mapped to the files and datasets
existing HDF5 users are familiar with.  This reduces the difficulty porting
applications to the new IO stack.

@@

Since HDF5~\cite{folk:2011:hdf5} offers an interface focused on files and the internal
data types, such as datasets, these concepts must be mapped onto the proposed
FFSIO data storage concepts.  This mapping is shown in Table~\ref{tab:mapping}

\begin{table}[ht]
%    \vspace{-0.10in}
    \centering
    \caption[HDF5 to FFSIO Mapping]{HDF5 Data Model to FFSIO Data Model Mapping}
    \bigskip
%    \vspace{-0.15in}
    \begin{tabular}{|r|r|}
\hline
HDF5 & FFSIO\\
\hline
file & container \\
dataset & array \\
group & key-value store \\
attribute(s) & key-value store \\
\hline
    \end{tabular}
    \label{tab:mapping}
\end{table}

In Section~\ref{sec:iod}, the FFSIO types are described in more detail.

@@

The core idea for IOD is to provide a way to manage IO load that is separate
from the compute nodes and the storage array. Communication intensive
activities, such as data rearrangement, can be moved to the IOD layer reducing
the number of participants and message count. The IOD has three main purposes.
First, the burst buffers work as a fast cache absorbing write operations that
then trickles out to the central storage array or pre-staging read operations.
It can also be used to retrieve objects from the central storage array for more
efficient read operations and offers data filtering to make client reads more
efficient.  Second, it offers the transaction mechanism for controlling data
set visibility and to manage faults that could expose an incomplete or corrupt
data set to users.  These transactions are local to the IOD layer until
persisted to the DAOS layer eliminating the need for burdening the persistent
storage with transient data.  Third, data processing operations can be placed
in the IOD.  These operations are intended to offer functionality like data
rearrangement and filtering prior to data reaching the central storage array.

@@

\textbf{Container}
As mentioned above, the concept of a container is similar to that of a file in
a traditional file system. However, rather than being in a directory structure,
each container essentially is stored in a hash-space allowing direct access to
any container without regard to the current organizational context of the file
system.  For example, there is no need to navigate a directory hierarchy to
name a particular container.

@@

\textbf{Key-Value Store}
This is the base type for the container. Since the container represents
something akin to HDF5 files, everything is stored within a hierarchical
namespace. The root namespace is represented by the base key-value store and
contains a list of all of the objects for this portion of the namespace as well
as additional key-value objects representing sub-groups for the hierarchy. Each
of those key-value store objects works identically.  Attributes are stored in a
key-value store object, but use the multi-dimensional array and blob objects to
store the values for the attributes.

\textbf{Multi-Dimensional Array}
By treating arrays as a special case separate from blobs, additional
opportunities are enabled. For example, by knowing that an object is an array,
proper slicing of that array onto IO nodes can be done without involving higher
levels of the IO stack.

\textbf{Blob}
All other data is stored as a stream-of-bytes without regard to the actual
data type.

\subsection{Multi-Dimensional Arrays}

For both IO performance and to aid in analysis and other data processing, the
multi-dimensional array object can be split across multiple IO nodes. Each
piece of this array is called a {\em shard}.  The idea of sharding is to store
a logically complete portion of a data set on a single storage target. This is
similar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
the data in the default or some other structured way as well as ``re-sharding''
based on application needs. For example, reordering the data so that a
different dimension is the ``fast'' dimension may greatly improve the
performance of a subsequent data analytics task. A common scenario where this
is useful is a Fortran code (column-major) writes data for a C code (row-major)
to analyze.  The IOD API supports the following sharding strategies:

\begin{itemize}
\item
  \textbf{contiguous}. Fixed chunking, distributed in a round-robin
  fashion across the IO nodes.
\item
  \textbf{chunked}. Same as above but with irregular (sparse) chunking.
\item
  \textbf{user-defined}. Either contiguous or chunked, but user specifies
  where to place each individual shard.
\end{itemize}

It is possible to request the transformation of an object's physical
layout to other formats resulting in multiple copies of the same objects in
multiple formats. Also, the user can pre-fetch objects from the storage cluster
into the IO nodes or read them directly from the storage cluster. At the
semantic level (HDF5), indices can be created for datasets resulting in being
able to read through an index instead of directly from the base array.

@@

The Distributed Application Object Storage (DAOS) layer serves as the
traditional parallel file system interface layer for the storage devices. This
is the consistent, global view of the underlying devices represented in this
stack by the VOSD layer.  This is the layer where the container/object model is
translated into the physical storage requirements dictated by the physical
storage underneath (the VOSD layer). The two key design elements of this layer
are the handling of epochs and the mapping of containers and objects to the
underlying storage.

@@

In the subsections that follow, we examine transactions and epochs, metadata
management, and function and analysis shipping.

\begin{figure*}[htbp!]
\centering
\vspace{-0.10in}
\subfigure[Write Hosts]{\label{fig:write-hosts}\includegraphics[width=\columnwidth]{images/write-hosts.png}}
\subfigure[Read Hosts]{\label{fig:read-hosts}\includegraphics[width=\columnwidth]{images/read-hosts.png}}
\vspace{-0.10in}
\caption{Functionality Demonstration Validation for Number of Hosts}
\label{fig:eval-hosts}
\vspace{-0.05in}
\end{figure*}

\begin{figure*}[htbp!]
\centering
\vspace{-0.10in}
\subfigure[Write Size]{\label{fig:write-size}\includegraphics[width=\columnwidth]{images/write-size.png}}
\subfigure[Read Size]{\label{fig:read-size}\includegraphics[width=\columnwidth]{images/read-size.png}}
\vspace{-0.10in}
\caption{Functionality Demonstration Validation for Data Sizes}
\label{fig:eval-size}
\vspace{-0.2in}
\end{figure*}

@@

In our previous work~\cite{lofstead:2014:ffsio-consistency}, we focused
strongly on not just exploring, but also critiquing using transactions, as
proposed, for consistency control. This section recaps some of that information
to give a more self-contained presentation.

@@

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
FFSIO uses a hybrid approach to metadata management that is halfway between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

@@

@@

All of the tests are performed on the Buffy Cray XC30-AC at LANL.  It consists
of 64 compute nodes each with dual, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz
and 64 GB RAM.  The interconnect is Cray Aries. There are 14 IO nodes
consisting of single socket, 8 core Intel Xeon ES-2670 CPUs at 2.6 GHz with 32
GB of RAM.  There are also a metadata, login, and 2 boot nodes.  The storage
array has disk and SSD partitions. One is a DDN Lustre system with 192 TB disk
usable with a minimum of 5 GB/sec performance. The rest are SSDs consisting of
an EMC flash array connected via FDR InfiniBand with 22 TB and 48 GB/sec write
performance.

@@

With the overall stack design a prototype implementation complete, refinements,
such as fault detection and recovery, can be designed and tested.  These and
other activities for phase two will ultimately generate what is likely to be
the next generation storage stack for extreme scale platforms.

\section{Acknowledgements}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.
